Dynamics Analysis: The Sound of Teaching Music 2

### 2020-03-20 script/RNotebook/dynamics.Rmd
updated comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Dynamics: Analysis: The Sound of Teaching Music 2'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of all analyses for dynamics.

@@

### Main predictions
- Participants will exaggerate forte and piano more while teaching.
- Participants will make a larger contrast between forte and piano.

@@

## Key Velocity (KV) - Dynamics
### 1. All - individual{#vel_dyn_ind_all}
- Average KV for each participant.

@@

##### Box plot

```{r vel_dyn_all_box, echo = FALSE}
p_vel_dyn_all_box <- ggplot(data = vel_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
p_vel_dyn_all_box
```

#### Statistics
##### Normality check
```{r normality1, echo = FALSE}
# Reduce unused factors
vel_dyn$Subcomponent <- factor(vel_dyn$Subcomponent)
vel_dyn_norm <- by(vel_dyn$Mean, list(vel_dyn$Condition, vel_dyn$Subcomponent), shapiro.test)
print(vel_dyn_norm)
```
  - Violate the normality assumption

##### Two-way ANOVA (ez)
```{r stats1, echo = FALSE}
vel_dyn_aov <- ezANOVA(
  data = df_vel_dyn
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_dyn_aov$ANOVA$F <- round(vel_dyn_aov$ANOVA$F, 4)
vel_dyn_aov$ANOVA$p <- round(vel_dyn_aov$ANOVA$p, 4)
print(vel_dyn_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats2, echo = FALSE}
vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
print(summary(vel_dyn_aov_2))
```

##### Post-hoc (direct comparison)
```{r post-hoc1, echo = FALSE}
vel_ttest_forte <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"], paired = TRUE)
vel_ttest_piano <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"], paired = TRUE)
# Effect size
vel_ttest_forte_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Forte"], vel_dyn$Condition[vel_dyn$Subcomponent == "Forte"], paired = TRUE)
vel_ttest_piano_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Piano"], vel_dyn$Condition[vel_dyn$Subcomponent == "Piano"], paired = TRUE)

print(vel_ttest_forte)
print(vel_ttest_piano)
print(vel_ttest_forte_cohend)
print(vel_ttest_piano_cohend)
```

##### Post-hoc (pairwise t test)
```{r post-hoc2, echo = FALSE}
# Create a group label for a post-hoc test
vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"] <- "Performing-Forte" 
vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"] <- "Performing-Piano"
vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"] <- "Teaching-Forte"
vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"] <- "Teaching-Piano"
# pairwise comparison
vel_dyn_pairwise <- pairwise.t.test(vel_dyn$Mean, vel_dyn$Group, paired = TRUE, p.adjust.method = "BH")
vel_dyn_pairwise$p.value <- round(vel_dyn_pairwise$p.value, 4)
print(vel_dyn_pairwise)
```

### 3. Trial by trial - individual{#vel_dyn_ind_trial}
- Average KV for each trial for each participant.

@@

#### Plot
##### Line graph
```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
p_vel_dyn_trial_forte <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Forte") +　scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_forte
```

@@

## Key Velocity Difference (KV-Diff) - Dynamics
### 1. All - individual{#vel_diff_dyn_ind_all}
- Average KV-Diff for each participant.

@@

### 2. All - group{#vel_diff_dyn_group_all}
- Average KV-Diff for each condition.
- Hypothesis: Participants will make a larger contrast between forte and piano.

@@

```{r vel_diff_dyn_all_bar,  echo = FALSE}
p_vel_diff_dyn_all <- ggplot(data = vel_diff_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_diff_dyn_all
```

##### Box plot

```{r vel_diff_dyn_all_box, echo = FALSE}
p_vel_diff_dyn_all_box <- ggplot(data = vel_diff_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_diff_dyn_all_box
```

#### Statistics
##### Normality check
```{r normality2, echo = FALSE}
# Reduce unused factors
vel_diff_dyn$Subcomponent <- factor(vel_diff_dyn$Subcomponent)
vel_diff_dyn_norm <- by(vel_diff_dyn$Mean, list(vel_diff_dyn$Condition, vel_diff_dyn$Subcomponent), shapiro.test)
print(vel_diff_dyn_norm)
```

##### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
vel_diff_dyn_aov <- ezANOVA(
  data = subset(df_vel_diff_dyn, df_vel_diff_dyn$Subcomponent == "FtoP" | df_vel_diff_dyn$Subcomponent == "PtoF")
  , dv = .(Diff)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_diff_dyn_aov$ANOVA$F <- round(vel_diff_dyn_aov$ANOVA$F, 4)
vel_diff_dyn_aov$ANOVA$p <- round(vel_diff_dyn_aov$ANOVA$p, 4)
print(vel_diff_dyn_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
vel_diff_dyn_aov_2 <- with(vel_diff_dyn,
                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
print(summary(vel_diff_dyn_aov_2))
```

##### Post-hoc (direct comparison)
```{r post-hoc3, echo = FALSE}
vel_diff_ttest_ftop <- t.test(vel_diff_dyn$Mean[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "FtoP"], vel_diff_dyn$Mean[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "FtoP"], paired = TRUE)
vel_diff_ttest_ptof <- t.test(vel_diff_dyn$Mean[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "PtoF"], vel_diff_dyn$Mean[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "PtoF"], paired = TRUE)
# Effect size
vel_diff_ttest_ftop_cohend <- cohen.d(vel_diff_dyn$Mean[vel_diff_dyn$Subcomponent == "FtoP"], vel_diff_dyn$Condition[vel_diff_dyn$Subcomponent == "FtoP"], paired = TRUE)
vel_diff_ttest_ptof_cohend <- cohen.d(vel_diff_dyn$Mean[vel_diff_dyn$Subcomponent == "PtoF"], vel_diff_dyn$Condition[vel_diff_dyn$Subcomponent == "PtoF"], paired = TRUE)

print(vel_diff_ttest_ftop)
print(vel_diff_ttest_ptof)
print(vel_diff_ttest_ftop_cohend)
print(vel_diff_ttest_ptof_cohend)
```

##### Post-hoc (pairwise t test)
```{r post-hoc4, echo = FALSE}
# Create a group label for a post-hoc test
vel_diff_dyn$Group[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "FtoP"] <- "Performing-FtoP" 
vel_diff_dyn$Group[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "PtoF"] <- "Performing-PtoF"
vel_diff_dyn$Group[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "FtoP"] <- "Teaching-FtoP"
vel_diff_dyn$Group[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "PtoF"] <- "Teaching-PtoF"
# pairwise comparison
vel_diff_dyn_pairwise <- pairwise.t.test(vel_diff_dyn$Mean, vel_diff_dyn$Group, paired = TRUE, p.adjust.method = "BH")
vel_diff_dyn_pairwise$p.value <- round(vel_diff_dyn_pairwise$p.value, 4)
print(vel_diff_dyn_pairwise)
```

### 3. Trial by trial - individual{#vel_diff_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

@@

#### Plot
##### Line graph
```{r vel_diff_dyn_trial_line_ftop, echo = FALSE, fig.height = 4}
p_vel_diff_dyn_trial_ftop <- ggplot(data = subset(vel_diff_dyn_trial, vel_diff_dyn_trial$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_diff_dyn_trial_ftop
```

@@

### 4. Trial by trial - group{#vel_diff_dyn_group_trial}
- Average KV-Diff for each trial for each condition.

@@

## Key Velocity (KV) - Articulation
### 1. All - individual{#vel_art_ind_all}
- Average KV for each participant.

@@

##### Box plot

```{r vel_art_all_box, echo = FALSE}
p_vel_art_all_box <- ggplot(data = vel_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation")
p_vel_art_all_box
```

#### Statistics
##### Normality check
```{r normality3, echo = FALSE}
# Reduce unused factors
vel_art$Subcomponent <- factor(vel_art$Subcomponent)
vel_art_norm <- by(vel_art$Mean, list(vel_art$Condition, vel_art$Subcomponent), shapiro.test)
print(vel_art_norm)
```

##### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
vel_art_aov <- ezANOVA(
  data = df_vel_art
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_art_aov$ANOVA$F <- round(vel_art_aov$ANOVA$F, 4)
vel_art_aov$ANOVA$p <- round(vel_art_aov$ANOVA$p, 4)
print(vel_art_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
print(summary(vel_art_aov_2))
```

### 3. Trial by trial - individual{#vel_art_ind_trial}
- Average KV for each trial for each participant.

@@

#### Plot
##### Line graph
```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 4}
p_vel_art_trial_legato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_legato
```

@@

## Key Velocity Difference (KV-Diff) - Articulation
### 1. All - individual{#vel_diff_art_ind_all}
- Average KV-Diff for each participant.

@@

### 2. All - group{#vel_diff_art_group_all}
- Average KV-Diff for each condition.

@@

##### Bar chart

@@

##### Box plot

```{r vel_diff_art_all_box, echo = FALSE}
p_vel_diff_art_all_box <- ggplot(data = vel_diff_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
p_vel_diff_art_all_box
```

#### Statistics
##### Normality check
```{r normality4, echo = FALSE}
# Reduce unused factors
vel_diff_art$Subcomponent <- factor(vel_diff_art$Subcomponent)
vel_diff_art_norm <- by(vel_diff_art$Mean, list(vel_diff_art$Condition, vel_diff_art$Subcomponent), shapiro.test)
print(vel_diff_art_norm)
```

##### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
vel_diff_art_aov <- ezANOVA(
  data = subset(df_vel_diff_art, df_vel_diff_art$Subcomponent == "LtoS" | df_vel_diff_art$Subcomponent == "StoL")
  , dv = .(Diff)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_diff_art_aov$ANOVA$F <- round(vel_diff_art_aov$ANOVA$F, 4)
vel_diff_art_aov$ANOVA$p <- round(vel_diff_art_aov$ANOVA$p, 4)
print(vel_diff_art_aov$ANOVA)
```

##### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
vel_diff_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_diff_art)
print(summary(vel_diff_art_aov_2))
```

### 3. Trial by trial - individual{#kot_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

@@

#### Plot
##### Line graph
```{r vel_diff_art_trial_line_ltos, echo = FALSE, fig.height = 4}
p_vel_diff_art_trial_ltos <- ggplot(data = subset(vel_diff_art_trial, vel_diff_art_trial$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_diff_art_trial_ltos
```

@@

### 4. Trial by trial - group{#vel_diff_art_group_trial}
- Average KV-Diff for each trial for each condition.

>>>>

---
title: 'Dynamics Analysis: The Sound of Teaching Music 2'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for dynamics.

- Last checked: 20/03/20

# Outline - Dynamics (Key Velocity)
## [Experimental Design](#design)

## Dynamics

### [KV](#vel-dynamics)
- Individual results
  + [All](#vel_dyn_ind_all)
  + [Trial by trial](#vel_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_dyn_group_all)
  + [Trial by trial](#vel_dyn_group_trial)

### [KV-Diff](#vel-dynamics-diff)
- Individual results
  + [All](#vel_diff_dyn_ind_all)
  + [Trial by trial](#vel_diff_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_diff_dyn_group_all)
  + [Trial by trial](#vel_diff_dyn_group_trial)

- [Sequence plot](#sequence_dyn)

## Articulation

### [KV](#vel-articulation)
- Individual results
  + [All](#vel_art_ind_all)
  + [Trial by trial](#vel_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_art_group_all)
  + [Trial by trial](#vel_art_group_trial)

### [KV-Diff](#vel-articulation-diff)
- Individual results
  + [All](#vel_diff_art_ind_all)
  + [Trial by trial](#vel_diff_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_diff_art_group_all)
  + [Trial by trial](#vel_diff_art_group_trial)

- [Sequence plot](#sequence_art)

@@

## Main predictions
- Participants will exaggerate forte and piano more while teaching.
- Participants will make a larger contrast between forte and piano.

@@

# Key Velocity (KV) - Dynamic{#vel-dynamics}
## 1. All - individual{#vel_dyn_ind_all}
- Average KV for each participant.

@@

#### Box plot
```{r vel_dyn_all_box, echo = FALSE}
p_vel_dyn_all_box <- ggplot(data = vel_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
p_vel_dyn_all_box
```

### Statistics
#### Normality check
```{r normality1, echo = FALSE}
# Reduce unused factors
vel_dyn$Subcomponent <- factor(vel_dyn$Subcomponent)
vel_dyn_norm <- by(vel_dyn$Mean, list(vel_dyn$Condition, vel_dyn$Subcomponent), shapiro.test)
print(vel_dyn_norm)
```
  - Violate the normality assumption

#### Two-way ANOVA (ez)
```{r stats1, echo = FALSE}
vel_dyn_aov <- ezANOVA(
  data = df_vel_dyn
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_dyn_aov$ANOVA$F <- round(vel_dyn_aov$ANOVA$F, 4)
vel_dyn_aov$ANOVA$p <- round(vel_dyn_aov$ANOVA$p, 4)
print(vel_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats2, echo = FALSE}
vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
print(summary(vel_dyn_aov_2))
```

#### Post-hoc (direct comparison)
```{r post-hoc1, echo = FALSE}
vel_ttest_forte <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"], paired = TRUE)
vel_ttest_piano <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"], paired = TRUE)

print(vel_ttest_forte)
print(vel_ttest_piano)
```

#### Effect size
```{r effect1, echo = FALSE}
# Effect size
vel_ttest_forte_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Forte"], vel_dyn$Condition[vel_dyn$Subcomponent == "Forte"], paired = TRUE)
vel_ttest_piano_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Piano"], vel_dyn$Condition[vel_dyn$Subcomponent == "Piano"], paired = TRUE)

print(vel_ttest_forte_cohend)
print(vel_ttest_piano_cohend)
```

#### Post-hoc (pairwise t test)
```{r post-hoc2, echo = FALSE}
# Create a group label for a post-hoc test
vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"] <- "Performing-Forte" 
vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"] <- "Performing-Piano"
vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"] <- "Teaching-Forte"
vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"] <- "Teaching-Piano"
# pairwise comparison
vel_dyn_pairwise <- pairwise.t.test(vel_dyn$Mean, vel_dyn$Group, paired = TRUE, p.adjust.method = "BH")
vel_dyn_pairwise$p.value <- round(vel_dyn_pairwise$p.value, 4)
print(vel_dyn_pairwise)
```

## 3. Trial by trial - individual{#vel_dyn_ind_trial}
- Average KV for each trial for each participant.

@@

### Plot
#### Line graph
```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
p_vel_dyn_trial_forte <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Forte") +　scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_forte
```

@@

# Key Velocity Difference (KV-Diff) - Dynamics{#vel-dynamics-diff}
## 1. All - individual{#vel_diff_dyn_ind_all}
- Average KV-Diff for each participant.

@@

## 2. All - group{#vel_diff_dyn_group_all}
- Average KV-Diff for each condition.
- Hypothesis: Participants will make a larger contrast between forte and piano.

@@

### Plot
#### Bar chart
```{r vel_diff_dyn_all_bar,  echo = FALSE}
p_vel_diff_dyn_all <- ggplot(data = vel_diff_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_diff_dyn_all
```

#### Box plot
```{r vel_diff_dyn_all_box, echo = FALSE}
p_vel_diff_dyn_all_box <- ggplot(data = vel_diff_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_diff_dyn_all_box
```

### Statistics
#### Normality check
```{r normality2, echo = FALSE}
# Reduce unused factors
vel_diff_dyn$Subcomponent <- factor(vel_diff_dyn$Subcomponent)
vel_diff_dyn_norm <- by(vel_diff_dyn$Mean, list(vel_diff_dyn$Condition, vel_diff_dyn$Subcomponent), shapiro.test)
print(vel_diff_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
vel_diff_dyn_aov <- ezANOVA(
  data = subset(df_vel_diff_dyn, df_vel_diff_dyn$Subcomponent == "FtoP" | df_vel_diff_dyn$Subcomponent == "PtoF")
  , dv = .(Diff)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_diff_dyn_aov$ANOVA$F <- round(vel_diff_dyn_aov$ANOVA$F, 4)
vel_diff_dyn_aov$ANOVA$p <- round(vel_diff_dyn_aov$ANOVA$p, 4)
print(vel_diff_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
vel_diff_dyn_aov_2 <- with(vel_diff_dyn,
                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
print(summary(vel_diff_dyn_aov_2))
```

#### Post-hoc (direct comparison)
```{r post-hoc3, echo = FALSE}
vel_diff_ttest_ftop <- t.test(vel_diff_dyn$Mean[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "FtoP"], vel_diff_dyn$Mean[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "FtoP"], paired = TRUE)
vel_diff_ttest_ptof <- t.test(vel_diff_dyn$Mean[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "PtoF"], vel_diff_dyn$Mean[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "PtoF"], paired = TRUE)

print(vel_diff_ttest_ftop)
print(vel_diff_ttest_ptof)
```

#### Effect size
```{r effect2, echo = FALSE}
# Effect size
vel_diff_ttest_ftop_cohend <- cohen.d(vel_diff_dyn$Mean[vel_diff_dyn$Subcomponent == "FtoP"], vel_diff_dyn$Condition[vel_diff_dyn$Subcomponent == "FtoP"], paired = TRUE)
vel_diff_ttest_ptof_cohend <- cohen.d(vel_diff_dyn$Mean[vel_diff_dyn$Subcomponent == "PtoF"], vel_diff_dyn$Condition[vel_diff_dyn$Subcomponent == "PtoF"], paired = TRUE)

print(vel_diff_ttest_ftop_cohend)
print(vel_diff_ttest_ptof_cohend)
```

#### Post-hoc (pairwise t test)
```{r post-hoc4, echo = FALSE}
# Create a group label for a post-hoc test
vel_diff_dyn$Group[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "FtoP"] <- "Performing-FtoP" 
vel_diff_dyn$Group[vel_diff_dyn$Condition == "performing" & vel_diff_dyn$Subcomponent == "PtoF"] <- "Performing-PtoF"
vel_diff_dyn$Group[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "FtoP"] <- "Teaching-FtoP"
vel_diff_dyn$Group[vel_diff_dyn$Condition == "teaching" & vel_diff_dyn$Subcomponent == "PtoF"] <- "Teaching-PtoF"
# pairwise comparison
vel_diff_dyn_pairwise <- pairwise.t.test(vel_diff_dyn$Mean, vel_diff_dyn$Group, paired = TRUE, p.adjust.method = "BH")
vel_diff_dyn_pairwise$p.value <- round(vel_diff_dyn_pairwise$p.value, 4)
print(vel_diff_dyn_pairwise)
```

## 3. Trial by trial - individual{#vel_diff_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

@@

### Plot
#### Line graph
```{r vel_diff_dyn_trial_line_ftop, echo = FALSE, fig.height = 4}
p_vel_diff_dyn_trial_ftop <- ggplot(data = subset(vel_diff_dyn_trial, vel_diff_dyn_trial$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_diff_dyn_trial_ftop
```

@@

## 4. Trial by trial - group{#vel_diff_dyn_group_trial}
- Average KV-Diff for each trial for each condition.

@@

# Key Velocity (KV) - Articulation{#vel-articulation}
## 1. All - individual{#vel_art_ind_all}
- Average KV for each participant.

@@

#### Box plot
```{r vel_art_all_box, echo = FALSE}
p_vel_art_all_box <- ggplot(data = vel_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation")
p_vel_art_all_box
```

### Statistics
#### Normality check
```{r normality3, echo = FALSE}
# Reduce unused factors
vel_art$Subcomponent <- factor(vel_art$Subcomponent)
vel_art_norm <- by(vel_art$Mean, list(vel_art$Condition, vel_art$Subcomponent), shapiro.test)
print(vel_art_norm)
```

#### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
vel_art_aov <- ezANOVA(
  data = df_vel_art
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_art_aov$ANOVA$F <- round(vel_art_aov$ANOVA$F, 4)
vel_art_aov$ANOVA$p <- round(vel_art_aov$ANOVA$p, 4)
print(vel_art_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
print(summary(vel_art_aov_2))
```

## 3. Trial by trial - individual{#vel_diff_art_ind_all}
- Average KV for each trial for each participant.

@@

### Plot
#### Line graph
```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 4}
p_vel_art_trial_legato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_legato
```

@@

# Key Velocity Difference (KV-Diff) - Articulation{#vel-articulation-diff}
## 1. All - individual{#vel_diff_art_ind_all}
- Average KV-Diff for each participant.

@@

## 2. All - group{#vel_diff_art_group_all}
- Average KV-Diff for each condition.

@@

### Plot
#### Bar chart
- Error bars show SEM of the mean

@@

#### Box plot
```{r vel_diff_art_all_box, echo = FALSE}
p_vel_diff_art_all_box <- ggplot(data = vel_diff_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
p_vel_diff_art_all_box
```

### Statistics
#### Normality check
```{r normality4, echo = FALSE}
# Reduce unused factors
vel_diff_art$Subcomponent <- factor(vel_diff_art$Subcomponent)
vel_diff_art_norm <- by(vel_diff_art$Mean, list(vel_diff_art$Condition, vel_diff_art$Subcomponent), shapiro.test)
print(vel_diff_art_norm)
```

#### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
vel_diff_art_aov <- ezANOVA(
  data = subset(df_vel_diff_art, df_vel_diff_art$Subcomponent == "LtoS" | df_vel_diff_art$Subcomponent == "StoL")
  , dv = .(Diff)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_diff_art_aov$ANOVA$F <- round(vel_diff_art_aov$ANOVA$F, 4)
vel_diff_art_aov$ANOVA$p <- round(vel_diff_art_aov$ANOVA$p, 4)
print(vel_diff_art_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
vel_diff_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_diff_art)
print(summary(vel_diff_art_aov_2))
```

## 3. Trial by trial - individual{#vel_art_ind_trial}
- Average KV-Diff for each trial for each participant.

@@

### Plot
#### Line graph
```{r vel_diff_art_trial_line_ltos, echo = FALSE, fig.height = 4}
p_vel_diff_art_trial_ltos <- ggplot(data = subset(vel_diff_art_trial, vel_diff_art_trial$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_diff_art_trial_ltos
```

@@

## 4. Trial by trial - group{#vel_diff_art_group_trial}
- Average KV-Diff for each trial for each condition.
